---
title: "Supplementary Materials to: A thorough evaluation of the Language Environment Analysis (LENA) system"
shorttitle        : "LENA evaluation"

author: 
  - name          : "Alejandrina Cristia"
    affiliation   : "1"
    corresponding : true    # Define only one corresponding author
    address       : "29, rue d’Ulm, 75005 Paris, France"
    email         : "alecristia@gmail.com"
  - name: Marvin Lavechin
    affiliation: "1"
  - name: Camila Scaff
    affiliation: "1"
  - name: Melanie Soderstrom
    affiliation: "2"
  - name: Caroline Rowland
    affiliation: "3"
  - name: Okko Räsänen
    affiliation: "4,5"
  - name: John Bunce
    affiliation: "2,"
  - name: Elika Bergelson
    affiliation: "6"

affiliation:
  - id            : "1"
    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d’études cognitives, ENS, EHESS, CNRS, PSL University"
  - id: "2"
    institution: "Department of Psychology, University of Manitoba, Canada"
  - id: "3"
    institution: "Max Planck Institute for Psycholinguistics, Netherlands"
  - id: "4"
    institution: "Unit of Computing Sciences, Tampere University, Finland"
  - id: "5"
    institution: "Department of Signal Processing and Acoustics, Aalto University, Finland"
  - id: "6"
    institution: "Psychology & Neuroscience, Duke University, Durham, North Carolina, USA"



abstract: >    
    This document provides the full results of analyses carried out in the main manuscript as well as other useful analyses.



bibliography      : ["lena_eval.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"

classoption: table    
header-includes:
  - \usepackage{array}
  - \usepackage{float}

output            : papaja::apa6_pdf

---


```{r setup, include=TRUE,echo=F,warning=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
options(kableExtra.latex.load_packages = FALSE)
require("papaja")
thisdir=c("../LENA_eval_201906/")
library(lme4)
library(sm)
library(scales)
library(ggpubr)
library(ggplot2)
library(car)
library(dplyr)
library(kableExtra)

dodiv=function(x) x/sum(x, na.rm=T)

```



```{r read-cm,echo=F}


ber=read.table(paste0(thisdir,"BER","_cm.txt"),header=T)
row=read.table(paste0(thisdir,"ROW","_cm.txt"),header=T)
sod=read.table(paste0(thisdir,"SOD","_cm.txt"),header=T)
war=read.table(paste0(thisdir,"WAR","_cm.txt"),header=T)
all=read.table(paste0(thisdir,"all","_cm.txt"),header=T)


#remove empty rows
ber=ber[rownames(ber)[rowSums(ber)>0],]
row=row[rownames(row)[rowSums(row)>0],]
war=war[rownames(war)[rowSums(war)>0],]
sod=sod[rownames(sod)[rowSums(sod)>0],]
all=sod[rownames(all)[rowSums(all)>0],]

sumref=rowSums(all)

```

```{r DIAER,echo=F}
read.csv(paste0(thisdir,"/diaer_gold_no_ele_no_ovl_lena_sil_no_tv_no_oln_report.csv"))->py
#dim(py) #873 clips
#summary(py)

#294 FA, MI, confusion are NA because no speech at all in the clip
pytot=subset(py,item=="TOTAL")
#other fixes and addtions
py=subset(py,item!="TOTAL")
py$false.alarm..[py$total==0 & py$false.alarm!=0]<-100
py$false.alarm..[py$total==0 & py$false.alarm==0]<-0
py$missed.detection..[py$total==0 ]<-0
py$missed.detection..[py$missed.detection==0 ]<-0
py$cor=substr(py$item,1,3)
py$cor[grep("[0-9]",py$cor)]<-"tsi"
py$cor=factor(py$cor)

py$child=substr(py$item,1,8)
py$child[py$cor=="tsi"]=substr(py$item[py$cor=="tsi"],1,3)

# add in age info
spreadsheet = read.csv(paste0(thisdir,"/ACLEW_list_of_corpora.csv"), header=TRUE, sep = ",")
spreadsheet$child=paste0(spreadsheet$labname,"_",ifelse(nchar(spreadsheet$aclew_id)==3,paste0("0",spreadsheet$aclew_id),spreadsheet$aclew_id))
spreadsheet = spreadsheet[,c("child","age_mo_round")]
colnames(spreadsheet) = c("child","age")
spreadsheet_tsi = read.csv(paste0(thisdir,"/anon_metadata.csv"), header=TRUE, sep = ",")
spreadsheet_tsi = spreadsheet_tsi[c("id","age_mo")]
colnames(spreadsheet_tsi) = c("child","age")
age_id = rbind(spreadsheet, spreadsheet_tsi)
py=merge(py,age_id,by.x="child",by.y="child",all.x=T)
```

## LENA^®^ classification accuracy: False alarms, misses, confusion 

No additional results.


```{r cf-dm,echo=F}

 read.csv("../LENA_eval_201906/diaer_LENA_sil.csv")->pysil
 read.csv("../LENA_eval_201906/diaer_LENA_far.csv")->pyfar
 read.csv("../LENA_eval_201906/diaer_gold_no_ele_LENA_sil_no_tv_report.csv")->pynotv

 pysil$type<-"pysil"
pyfar$type<-"pyfar"
pynotv$type<-"pynotv"

clean<-function(py){
  py[py$item!="Total",]->py
  py$cor=substr(py$item,1,1)
  py$cor[py$cor=="C"] ="T"
  py
}

allpy= rbind(pysil,pyfar,pynotv)

clean(allpy)->allpy 

medians=aggregate(allpy[,c("false.alarm..","missed.detection..","confusion..")],by=list(allpy$type),median,na.rm=T)

colnames(medians)[1]<-c("type")
medians[2:4]<-round(medians[2:4])
medians$der=rowSums(medians[2:4])
```


## LENA^®^ classification accuracy: Precision and recall 

No additional results.

### LENA^®^ classification accuracy: Precision

No additional results.


### LENA^®^ classification accuracy: Recall 
No additional results.


## Derived metrics

### Child Vocalization Counts (CVC) accuracy 

```{r cvc,echo=F}
read.table(paste0(thisdir,"/cvtc.txt"),header=T)->cvtc



cor.cvc.all= cor.test(cvtc$CVC_n,cvtc$CVC_gold)
cor.cvc.noZeros= cor.test(cvtc$CVC_n[cvtc$CVC_n>0 & cvtc$CVC_gold>0],cvtc$CVC_gold[cvtc$CVC_n>0 & cvtc$CVC_gold>0])

```

No additional results.


### Conversational Turn Counts (CTC) accuracy 

No additional results.

### Adult Word Counts accuracy 

```{r awc, echo=F}
read.table("../LENA_AWC_rel_v1_June.txt")->awc
colnames(awc)<-c("filename","gold","LENA")
gsub("LUC","ROW",awc$filename)->awc$filename

awc$cor=substr(awc$filename,1,3)
awc$cor[grep("[0-9]",awc$cor)]<-"tsi"
awc$cor=factor(awc$cor)

awc$child=substr(awc$filename,1,8)
awc$child[awc$cor=="tsi"]=substr(awc$X[awc$cor=="tsi"],1,3)


merge(awc,age_id,by="child")->awc


cor.awc.all= cor.test(awc$gold,awc$LENA)
cor.awc.noZeros= cor.test(awc$gold[awc$gold>0 & awc$LENA>0],awc$LENA[awc$gold>0 & awc$LENA>0])

# ER against human
aer=(awc$LENA-awc$gold)
aer_no_zeros=(awc$LENA[awc$LENA>0 & awc$gold > 0]-awc$gold[awc$LENA>0 & awc$gold > 0])

no_human_zeros=subset(awc,gold>0)
rer=(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100

arer=abs(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100

aer_awc <-(awc$LENA-awc$gold)
aer_nozeros_awc <-(awc$LENA[awc$LENA>0 & awc$gold > 0]-awc$gold[awc$LENA>0 & awc$gold > 0])
```

```{r awc-fig, ig.pos = "T!", fig.width=3.5, fig.height=3.5, fig.cap="Adult Word Counts according to LENA (x axis) and humans (y axis). Each point represents the AWC totaled within a clip. The solid line corresponds to a linear regression fit to data from all clips; the dashed line corresponds to an analysis excluding clips where both the human and LENA^®^ said there were no adult words."}

plot(gold~LENA,data=awc,pch=20,main="AWC",col=alpha("orange",.2),xlab="LENA AWC",ylab="Human AWC")
abline(lm(gold~LENA,data=awc))
abline(lm(gold~LENA,data=awc,subset=c(awc$gold>0 & awc$LENA>0)),lty=2)
```

No additional results.



## Effects of age and differences across corpora 

We report on full model results.

### Diarization error rate

```{r lmer-dm, echo=F}
read.csv(paste0(thisdir,"/diaer_gold_no_ele_no_ovl_lena_sil_no_tv_no_oln_report.csv"))->py

#other fixes and addtions
py=subset(py,item!="TOTAL")
py$cor=substr(py$item,1,3)
py$cor[grep("[0-9]",py$cor)]<-"tsi"
py$cor=factor(py$cor)

py$child=substr(py$item,1,8)
py$child[py$cor=="tsi"]=substr(py$item[py$cor=="tsi"],1,3)

py$false.alarm..[py$total==0 & py$false.alarm!=0]<-100
py$false.alarm..[py$total==0 & py$false.alarm==0]<-0

py$missed.detection..[py$total==0 ]<-0
py$missed.detection..[py$missed.detection==0 ]<-0

# add in age info
spreadsheet = read.csv(paste0(thisdir,"/ACLEW_list_of_corpora.csv"), header=TRUE, sep = ",")
spreadsheet$child=paste0(spreadsheet$labname,"_",ifelse(nchar(spreadsheet$aclew_id)==3,paste0("0",spreadsheet$aclew_id),spreadsheet$aclew_id))
spreadsheet = spreadsheet[,c("child","age_mo_round")]
colnames(spreadsheet) = c("child","age")
spreadsheet_tsi = read.csv(paste0(thisdir,"/anon_metadata.csv"), header=TRUE, sep = ",")
spreadsheet_tsi = spreadsheet_tsi[c("id","age_mo")]
colnames(spreadsheet_tsi) = c("child","age")
age_id = rbind(spreadsheet, spreadsheet_tsi)
py=merge(py,age_id,by.x="child",by.y="child",all.x=T) #one row is lost... **ATTENTION BUG HERE**

mydv=c("false.alarm..","missed.detection..","confusion..")

mymodels<- list()
myANOVAs<- list()
anovatab=matrix(nrow=4,ncol=3)
colnames(anovatab)<-mydv
for(dv in mydv){
    print(paste("The following model corresponds to",dv))
    mymodels[[dv]]<-lmer(py[,dv]~cor*age+(1|child),data=py)
  myANOVAs[[dv]]<-Anova(mymodels[[dv]],type="III")
  print(summary(mymodels[[dv]]))
  print(myANOVAs[[dv]])
  mycols=cbind(round(myANOVAs[[dv]]$Chisq,2),round(myANOVAs[[dv]]$Df),round(myANOVAs[[dv]]["Pr(>Chisq)"],2))
  anovatab[,dv]<-paste0(mycols[,1]," (",mycols[,2],")",ifelse(mycols[,3]<.05," *",""))
}
rownames(anovatab)<-rownames(mycols)

```

### CVC

```{r cvc-age, echo=F}
read.table(paste0(thisdir,"/cvtc.txt"),header=T)->cvtc

cvtc$cor=substr(cvtc$filename,1,3)
cvtc$cor[substr(cvtc$filename,1,1)=="C"]<-"tsi"
cvtc$cor=factor(cvtc$cor)

cvtc$child=substr(cvtc$filename,1,8)
cvtc$child[cvtc$cor=="tsi"]<-substr(cvtc$filename[cvtc$cor=="tsi"],1,3)

merge(cvtc,age_id,by="child")->cvtc

mymodel<-lmer(CVC_gold~CVC_n*age*cor + (1|child), data=cvtc)
summary(mymodel)
Anova(mymodel,type="III")

```


There is a 3-way interaction between age, corpus, and the predictive value of the LENA^®^ system's  counts with respect to the gold counts. To investigate this we fit the same regression within each corpus.


```{r 3way,echo=F}


mymodels<- list()
myANOVAs<- list()
for(thiscor in levels(cvtc$cor)){
    print(paste("The following model corresponds to",thiscor))
 mymodels[[thiscor]]<-lmer(CVC_gold~CVC_n*age + (1|child), data=cvtc,subset=c(cor==thiscor))
 myANOVAs[[thiscor]]<-Anova(mymodels[[thiscor]],type="III")
  print(summary(mymodels[[thiscor]]))
  print(myANOVAs[[thiscor]])
}

```


### CTC

```{r ctc-age, echo=F}

mymodel<-lmer(CTC_gold~CTC_n*age*cor + (1|child), data=cvtc)
summary(mymodel)
Anova(mymodel,type="III")

```


### AWC

```{r awc-age, echo=F}

mymodel<-lmer(gold~LENA*age*cor + (1|child), data=awc)
summary(mymodel)
Anova(mymodel,type="III")

```


There is a 3-way interaction between age, corpus, and the predictive value of the LENA^®^ system's  counts with respect to the gold counts. To investigate this we fit the same regression within each corpus.


```{r 3way,echo=F}



mymodels<- list()
myANOVAs<- list()
for(thiscor in levels(awc$cor)){
  print(paste("the model coming up corresponds to",thiscor))
 mymodels[[thiscor]]<-lmer(gold~LENA*age + (1|child), data=awc,subset=c(cor==thiscor))
 myANOVAs[[thiscor]]<-Anova(mymodels[[thiscor]],type="III")
 print(mymodels[[thiscor]])
 print(myANOVAs[[thiscor]])
}

```



\newpage



# References

\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}
