---
title: "Supplementary Materials to: A thorough evaluation of the Language Environment Analysis (LENA) system"
shorttitle        : "LENA evaluation"

author: 
  - name          : "Alejandrina Cristia"
    affiliation   : "1"
    corresponding : true    # Define only one corresponding author
    address       : "29, rue d’Ulm, 75005 Paris, France"
    email         : "alecristia@gmail.com"
  - name: Marvin Lavechin
    affiliation: "1"
  - name: Camila Scaff
    affiliation: "1"
  - name: Melanie Soderstrom
    affiliation: "2"
  - name: Caroline Rowland
    affiliation: "3"
  - name: Okko Räsänen
    affiliation: "4,5"
  - name: John Bunce
    affiliation: "2,"
  - name: Elika Bergelson
    affiliation: "6"

affiliation:
  - id            : "1"
    institution   : "Laboratoire de Sciences Cognitives et de Psycholinguistique, Département d’études cognitives, ENS, EHESS, CNRS, PSL University"
  - id: "2"
    institution: "Department of Psychology, University of Manitoba, Canada"
  - id: "3"
    institution: "Max Planck Institute for Psycholinguistics, Netherlands"
  - id: "4"
    institution: "Unit of Computing Sciences, Tampere University, Finland"
  - id: "5"
    institution: "Department of Signal Processing and Acoustics, Aalto University, Finland"
  - id: "6"
    institution: "Psychology & Neuroscience, Duke University, Durham, North Carolina, USA"



abstract: >    
    This document provides the full results of analyses carried out in the main manuscript as well as other useful analyses.



bibliography      : ["lena_eval.bib"]

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"

classoption: table    
header-includes:
  - \usepackage{array}
  - \usepackage{float}

output            : papaja::apa6_pdf

---


```{r setup, include=TRUE,echo=F,warning=FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
options(kableExtra.latex.load_packages = FALSE)
require("papaja")
evaldir=c("../evaluations/")
library(lme4)
library(sm)
library(scales)
library(ggpubr)
library(ggplot2)
library(car)
library(dplyr)
library(kableExtra)

dodiv=function(x) x/sum(x, na.rm=T)

```



```{r readin,echo=F}
source("readin_cm.R")
source("readin_ider.R")
source("readin_derived.R")

```

## LENA^®^ classification accuracy: False alarms, misses, confusion 

No additional results.


## LENA^®^ classification accuracy: Precision and recall 

No additional results.

### LENA^®^ classification accuracy: Precision

No additional results.


### LENA^®^ classification accuracy: Recall 
No additional results.

#### LENA^®^ classification accuracy: Agreement using Cohen's kappa
No additional results.

## Derived metrics

```{r cvc,echo=F}

#create matrices that will hold all results
allpearsons=matrix(nrow=6,ncol=4) #r's
rownames(allpearsons)<-c("CVC","CHN c","CTC","\" all CHI","AWC","\" no French")
colnames(allpearsons)<-c("N clips","r all","N","r")

aller=matrix(nrow=6,ncol=12) #error rates
rownames(aller)<-c("CVC","CHN c","CTC","\" all CHI","AWC","\" no French")
colnames(aller)<-c("AER","min","max","AER-0","min","max","RER","min1","max1","ARER","min2","max2")

#CVC, same as in main paper
cor.cvc.all= cor.test(cvtc$lena_CV_count,cvtc$gold_CV_count)
cor.cvc.noZeros= cor.test(cvtc$lena_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count>0],cvtc$gold_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count>0])

allpearsons["CVC","N clips"]<-round(cor.cvc.all$parameter)
allpearsons["CVC","r all"]<-round(cor.cvc.all$estimate,3)
allpearsons["CVC","N"]<-round(cor.cvc.noZeros$parameter)
allpearsons["CVC","r"]<-round(cor.cvc.noZeros$estimate,3)

aer=(cvtc$lena_CV_count-cvtc$gold_CV_count)
aer_no_zeros=(cvtc$lena_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count > 0]-cvtc$gold_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count > 0])
no_human_zeros=cvtc[ cvtc$gold_CV_count>0,]
rer=(no_human_zeros$lena_CV_count-no_human_zeros$gold_CV_count)/no_human_zeros$gold_CV_count*100
arer=abs(no_human_zeros$lena_CV_count-no_human_zeros$gold_CV_count)/no_human_zeros$gold_CV_count*100
aer_cvc <-(cvtc$lena_CV_count-cvtc$gold_CV_count)
aer_nozeros_cvc <-(cvtc$lena_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count > 0]-cvtc$gold_CV_count[cvtc$lena_CV_count>0 & cvtc$gold_CV_count > 0])


aller["CVC",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )


#CHN counts, added here
cor.chnc.all= cor.test(cvtc$lena_CH_count,cvtc$gold_CH_count)
cor.chnc.noZeros= cor.test(cvtc$lena_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count>0],cvtc$gold_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count>0])

allpearsons["CHN c","N clips"]<-round(cor.cvc.all$parameter)
allpearsons["CHN c","r all"]<-round(cor.cvc.all$estimate,3)
allpearsons["CHN c","N"]<-round(cor.cvc.noZeros$parameter)
allpearsons["CHN c","r"]<-round(cor.cvc.noZeros$estimate,3)

aer=(cvtc$lena_CH_count-cvtc$gold_CH_count)
aer_no_zeros=(cvtc$lena_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count > 0]-cvtc$gold_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count > 0])
no_human_zeros=cvtc[ cvtc$gold_CH_count>0,]
rer=(no_human_zeros$lena_CH_count-no_human_zeros$gold_CH_count)/no_human_zeros$gold_CH_count*100
arer=abs(no_human_zeros$lena_CH_count-no_human_zeros$gold_CH_count)/no_human_zeros$gold_CH_count*100
aer_cvc <-(cvtc$lena_CH_count-cvtc$gold_CH_count)
aer_nozeros_cvc <-(cvtc$lena_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count > 0]-cvtc$gold_CH_count[cvtc$lena_CH_count>0 & cvtc$gold_CH_count > 0])


aller["CHN c",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )

#CTC as in main paper
cor.ctc.all=cor.test(cvtc$lena_CTC_count,cvtc$gold_CTC_count)

cor.ctc.noZeros=cor.test(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count>0],cvtc$gold_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count>0])


allpearsons["CTC","N clips"]<-round(cor.ctc.all$parameter)
allpearsons["CTC","r all"]<-round(cor.ctc.all$estimate,3)
allpearsons["CTC","N"]<-round(cor.ctc.noZeros$parameter)
allpearsons["CTC","r"]<-round(cor.ctc.noZeros$estimate,3)

aer=(cvtc$lena_CTC_count-cvtc$gold_CTC_count)
aer_no_zeros=(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count > 0]-cvtc$gold_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count > 0])
no_human_zeros=cvtc[ cvtc$gold_CTC_count>0,]
rer=(no_human_zeros$lena_CTC_count-no_human_zeros$gold_CTC_count)/no_human_zeros$gold_CTC_count*100
arer=abs(no_human_zeros$lena_CTC_count-no_human_zeros$gold_CTC_count)/no_human_zeros$gold_CTC_count*100
aer_ctc <-(cvtc$lena_CTC_count-cvtc$gold_CTC_count)
aer_nozeros_ctc <-(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count > 0]-cvtc$gold_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count > 0])

aller["CTC",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )

#CTC counting all vocs in human
cor.ctc.all=cor.test(cvtc$lena_CTC_count,cvtc$gold_CTC_count_all)

cor.ctc.noZeros=cor.test(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all>0],cvtc$gold_CTC_count_all[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all>0])


allpearsons["\" all CHI","N clips"]<-round(cor.ctc.all$parameter)
allpearsons["\" all CHI","r all"]<-round(cor.ctc.all$estimate,3)
allpearsons["\" all CHI","N"]<-round(cor.ctc.noZeros$parameter)
allpearsons["\" all CHI","r"]<-round(cor.ctc.noZeros$estimate,3)

aer=(cvtc$lena_CTC_count-cvtc$gold_CTC_count_all)
aer_no_zeros=(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all > 0]-cvtc$gold_CTC_count_all[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all > 0])
no_human_zeros=cvtc[ cvtc$gold_CTC_count_all>0,]
rer=(no_human_zeros$lena_CTC_count-no_human_zeros$gold_CTC_count_all)/no_human_zeros$gold_CTC_count_all*100
arer=abs(no_human_zeros$lena_CTC_count-no_human_zeros$gold_CTC_count_all)/no_human_zeros$gold_CTC_count_all*100
aer_ctc <-(cvtc$lena_CTC_count-cvtc$gold_CTC_count_all)
aer_nozeros_ctc <-(cvtc$lena_CTC_count[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all > 0]-cvtc$gold_CTC_count_all[cvtc$lena_CTC_count>0 & cvtc$gold_CTC_count_all > 0])

aller["\" all CHI",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )

#AWC as in main paper
cor.awc.all= cor.test(awc$gold,awc$LENA)
cor.awc.noZeros= cor.test(awc$gold[awc$gold>0 & awc$LENA>0],awc$LENA[awc$gold>0 & awc$LENA>0])


allpearsons["AWC","N clips"]<-round(cor.awc.all$parameter)
allpearsons["AWC","r all"]<-round(cor.awc.all$estimate,3)
allpearsons["AWC","N"]<-round(cor.awc.noZeros$parameter)
allpearsons["AWC","r"]<-round(cor.awc.noZeros$estimate,3)

aer=(awc$LENA-awc$gold)
aer_no_zeros=(awc$LENA[awc$LENA>0 & awc$gold > 0]-awc$gold[awc$LENA>0 & awc$gold > 0])
no_human_zeros=subset(awc,gold>0)
rer=(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100
arer=abs(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100
aer_awc <-(awc$LENA-awc$gold)
aer_nozeros_awc <-(awc$LENA[awc$LENA>0 & awc$gold > 0]-awc$gold[awc$LENA>0 & awc$gold > 0])

aller["AWC",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )

#AWC without French child
noFrench=awc[awc$child!="SOD_1499",]
cor.noFrench.all= cor.test(noFrench$gold,noFrench$LENA)
cor.noFrench.noZeros= cor.test(noFrench$gold[noFrench$gold>0 & noFrench$LENA>0],noFrench$LENA[noFrench$gold>0 & noFrench$LENA>0])


allpearsons["\" no French","N clips"]<-round(cor.noFrench.all$parameter)
allpearsons["\" no French","r all"]<-round(cor.noFrench.all$estimate,3)
allpearsons["\" no French","N"]<-round(cor.noFrench.noZeros$parameter)
allpearsons["\" no French","r"]<-round(cor.noFrench.noZeros$estimate,3)

aer=(noFrench$LENA-noFrench$gold)
aer_no_zeros=(noFrench$LENA[noFrench$LENA>0 & noFrench$gold > 0]-noFrench$gold[noFrench$LENA>0 & noFrench$gold > 0])
no_human_zeros=subset(noFrench,gold>0)
rer=(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100
arer=abs(no_human_zeros$LENA-no_human_zeros$gold)/no_human_zeros$gold*100
aer_noFrench <-(noFrench$LENA-noFrench$gold)
aer_nozeros_noFrench <-(noFrench$LENA[noFrench$LENA>0 & noFrench$gold > 0]-noFrench$gold[noFrench$LENA>0 & noFrench$gold > 0])

aller["\" no French",]<-cbind(mean(aer),min(aer),max(aer),
                     mean(aer_no_zeros),min(aer_no_zeros),max(aer_no_zeros),
                     mean(rer),min(rer),max(rer),
                     mean(arer),min(arer),max(arer)
                     )

```

### Child Vocalization Counts (CVC) accuracy 

In the present version of the main paper, we followed the LENA^®^ definitions of what "counts" as a child vocalization for LENA^®^-reported CVC: Any vocalization that contains an utterance, i.e., a linguistic segment. Thus, child vocalizations consisting purely of vegetative/fix subsegments, or of cries, does not count towards CVC. In a previous version, however, we were counting any vocalization - regardless of its content. It turns out that accuracy is slightly better with the latter definition - which we will call CHN counts (so as to avoid a confusion with LENA^®^-reported Child Vocalization Counts).


```{r CHN-count-fig, ig.pos = "T!", fig.width=3.5, fig.height=3.5, fig.cap="Key Child Segment counts recalculated from LENA reports (x axis; CHN counts, counting CHN segments regardless of whether they contain some linguistic stretches or not) and humans (y axis; CHI counts, counting segments attributed to the key child regardless of whether they were classified as linguistic or not). Each point represents the child segments totaled within a clip. The solid line corresponds to a linear regression fit to data from all clips; the dashed line corresponds to an analysis excluding clips where both the human and LENA^®^  found zero child segments. The x and y ranges have been adjusted to be equal regardless of the data distribution."}
myrange=range(cvtc[,c("gold_CH_count","lena_CH_count")],na.rm=T)
plot(gold_CH_count~lena_CH_count,data=cvtc,pch=20,col=alpha("black",.2),xlab="LENA CHN counts",ylab="Human CHI counts",xlim=myrange,ylim=myrange)
abline(lm(gold_CH_count~lena_CH_count,data=cvtc))
abline(lm(gold_CH_count~lena_CH_count,data=cvtc,subset=c(cvtc$lena_CH_count>0 & cvtc$gold_CH_count>0)),lty=2)

```



```{r tabrs, echo=F}

apa_table(allpearsons,caption="For each measure (CVC, CHN counts, CTC, \" all CHI, AWC, \" no French -- see main text for details), N all stands for the total number of clips considered in the general correlation analysis and resulting Pearson r coefficient; N stands for the number of non-null clips (i.e., having some vocalizations, turns, and adult words respectively) as well as Pearson r in the analyses restricted to the non-null clips.",digits=c(0,0,3,0,3))
```


```{r tabrers, echo=F}

apa_table(aller,caption="For each measure (CVC, CHN counts, CTC, \" all CHI, AWC, \" no French -- see main text for details), AER is the absolute error rate, AER-0 the same excluding clips with counts of zero according to either the human or the system, RER is the relative error rate and ARER the absolute relative error rate (the latter two exclude clips where the human count is zero). ",digits=0)
```


### Conversational Turn Counts (CTC) accuracy 

As with CVC, LENA^®^ counts infant-adult turns only if a child vocalization contains some linguistic material. In the main text, this is the analysis we report on, but we thought it informative to include in the present supplementary materials a redefinition whereby turns are counted regardless of whether the CHN segment contains some linguistic material or not. For instance, an infant cry followed by a child phrase would be counted in the latter definition but not the former. 


```{r CTC-all-count-fig, ig.pos = "T!", fig.width=3.5, fig.height=3.5, fig.cap="Conversational turns between Key Child and an adult recalculated from LENA reports and counting all CHN vocalizations, even if they do not contain any linguistic material (x axis) and humans (y axis; similarly counting all segments attributed to the key child regardless of whether they were classified as linguistic or not). Each point represents the turns totaled within a clip. The solid line corresponds to a linear regression fit to data from all clips; the dashed line corresponds to an analysis excluding clips where both the human and LENA^®^  found zero turns. The x and y ranges have been adjusted to be equal regardless of the data distribution."}
myrange=range(cvtc[,c("gold_CH_count","lena_CH_count")],na.rm=T)
plot(gold_CH_count~lena_CH_count,data=cvtc,pch=20,col=alpha("black",.2),xlab="LENA CHN counts",ylab="Human CHI counts",xlim=myrange,ylim=myrange)
abline(lm(gold_CH_count~lena_CH_count,data=cvtc))
abline(lm(gold_CH_count~lena_CH_count,data=cvtc,subset=c(cvtc$lena_CH_count>0 & cvtc$gold_CH_count>0)),lty=2)

```



### Adult Word Counts accuracy 

One of the children in the corpus was in a French-speaking environment. The following Figure shows results for AWC excluding the 15 data points corresponding to this child.

```{r awc-fig, ig.pos = "T!", fig.width=3.5, fig.height=3.5, fig.cap="Adult Word Counts according to LENA (x axis) and humans (y axis) excluding one child in a French-speaking environment. Each point represents the AWC totaled within a clip. The solid line corresponds to a linear regression fit to data from all clips; the dashed line corresponds to an analysis excluding clips where both the human and LENA^®^ said there were no adult words."}

plot(gold~LENA,data=noFrench,pch=20,main="AWC",col=alpha("orange",.2),xlab="LENA AWC",ylab="Human AWC")
abline(lm(gold~LENA,data=noFrench))
abline(lm(gold~LENA,data=noFrench,subset=c(noFrench$gold>0 & noFrench$LENA>0)),lty=2)
```
\newpage


## Effects of age and differences across corpora 

Here we report on full model results.

### Identification error rate

```{r lmer-dm, echo=F}
mydv=c("false.alarm..","missed.detection..","confusion..")
mymodels<- list()
myANOVAs<- list()
anovatab=matrix(nrow=4,ncol=3)
colnames(anovatab)<-mydv
for(dv in mydv){
    print(paste("The following model corresponds to",dv))
    mymodels[[dv]]<-lmer(py[,dv]~cor*age+(1|child),data=py)
  myANOVAs[[dv]]<-Anova(mymodels[[dv]],type="III")
  print(summary(mymodels[[dv]]))
  print(myANOVAs[[dv]])
  mycols=cbind(round(myANOVAs[[dv]]$Chisq,2),round(myANOVAs[[dv]]$Df),round(myANOVAs[[dv]]["Pr(>Chisq)"],2))
  anovatab[,dv]<-paste0(mycols[,1]," (",mycols[,2],")",ifelse(mycols[,3]<.05," *",""))
}
rownames(anovatab)<-rownames(mycols)

```

There is a 2-way interaction between age and corpus for confusion. To investigate this we fit the same regression within each corpus.

```{r confusion-followup, echo=F}
mymodels2<- list()
myANOVAs2<- list()
for(thiscor in levels(py$cor)){ 
  print(thiscor)
 mymodels2[[thiscor]]<-lmer(py[,"confusion.."]~age + (1|child), data=py,subset=c(cor==thiscor))
 print(mymodels2[[thiscor]])
 myANOVAs2[[thiscor]]<-Anova(mymodels2[[thiscor]],type="III")
 print(myANOVAs2[[thiscor]])
}

```


\newpage

### CVC

```{r cvc-age, echo=F}

mymodel<-lmer(gold_CV_count~lena_CV_count*age*cor + (1|child), data=cvtc)
summary(mymodel)
Anova(mymodel,type="III")

```


There is a 3-way interaction between age, corpus, and the predictive value of the LENA^®^ system's  counts with respect to the gold counts. To investigate this we fit the same regression within each corpus.


```{r 3way-cvc,echo=F}


mymodels<- list()
myANOVAs<- list()
for(thiscor in levels(cvtc$cor)){
    print(paste("The following model corresponds to",thiscor))
 mymodels[[thiscor]]<-lmer(gold_CV_count~lena_CV_count*age + (1|child), data=cvtc,subset=c(cor==thiscor))
 myANOVAs[[thiscor]]<-Anova(mymodels[[thiscor]],type="III")
  print(summary(mymodels[[thiscor]]))
  print(myANOVAs[[thiscor]])
}

```

\newpage

### CTC

```{r ctc-age, echo=F}

mymodel<-lmer(gold_CTC_count~lena_CTC_count*age*cor + (1|child), data=cvtc)
summary(mymodel)
Anova(mymodel,type="III")

```
\newpage


### AWC

```{r awc-age, echo=F}

mymodel<-lmer(gold~LENA*age*cor + (1|child), data=awc)
summary(mymodel)
Anova(mymodel,type="III")

```


There is a 3-way interaction between age, corpus, and the predictive value of the LENA^®^ system's  counts with respect to the gold counts. To investigate this we fit the same regression within each corpus.


```{r 3way-awc,echo=F}



mymodels<- list()
myANOVAs<- list()
for(thiscor in levels(awc$cor)){
  print(paste("the model coming up corresponds to",thiscor))
 mymodels[[thiscor]]<-lmer(gold~LENA*age + (1|child), data=awc,subset=c(cor==thiscor))
 myANOVAs[[thiscor]]<-Anova(mymodels[[thiscor]],type="III")
 print(mymodels[[thiscor]])
 print(myANOVAs[[thiscor]])
}

```





