---
title: "DIHARD_comparison"
author: "EB"
date: "1/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(skimr)
```

Looking at the DIHARD2 data that Neville sent along to Alex to add to eval.
(I wasn't sure if this data was okay for public facing purposes so i just put it in a dropbox folder, but can move to github, just lmk. I.e. this won't knit for you if you don't also put the data from neville in that path)

Just looking at track 1 and 2 since it was single channel 

* Track 1 – single channel audio using reference SAD 
* Track 2 – single channel audio using system SAD




```{r read_in_data}

track1 <- read_tsv("~/Dropbox/dihard_lena_eval/dihard2_scoring/data/track1/scores_by_domain.tsv") %>% 
  filter(Domain=="CHILD" & IsFinal==T) %>% 
  mutate(track = "track1")
track2 <-read_tsv("~/Dropbox/dihard_lena_eval/dihard2_scoring/data/track2/scores_by_domain.tsv") %>% 
  filter(Domain=="CHILD" & IsFinal==T) %>% 
  mutate(track = "track2")

track1_2 <- bind_rows(track1, track2)
#View(track1)
#View(track2)
```

# Initial look

* any other filtering i should be doing?
```{r initial_look}
summary(track1_2)
glimpse(track1_2)
skim(track1_2)
```

## Boxplot 
but the value's here don't look exactly the same as his CHILD boxplot in this folder: is2019_dihard2_figures
for DIHARD2 track 1 or track 2 DER

* maybe he used different filtering on the data?
* maybe i shouldn't allow multiples per User/team?
```{r boxplot_data}
ggplot(track1_2, aes(Domain, DER))+geom_boxplot()+
  facet_wrap(~track)

```

## Summary descriptives
(mean, median, min, max DER)

```{r summary_stats}
track1_2 %>% 
  summarise(mean(DER),
            median(DER),
            min(DER),
            max(DER)
            )

track1_2 %>% 
  group_by(track) %>% 
  summarise(mean(DER),
            median(DER),
            min(DER),
            max(DER)
            )


```

